\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-04-26 paper/paper.tex

\subsection{Background}
Products can frequently fail due to more than one reason.  For example, there are many parts in a washing machine that can break causing the whole machine to fail.  Decorative tree lighting can fail if one bulb in the chain burns out.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \cite{nelson}.  A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.  Thus, the overall reliability of a unit can be modeled using the product rule across all $s$ risks.  The GFLP model is a special case of the $s$ independent competing risks model with 2 modes of failure: defects and wearout.  Sometimes the cause of the failure is know; but more often the observations do not contain information about the failure mode.  The GFLP model was first introduced to describe the early and late failure modes of CB radios \cite{chan}.  No failure information was available for this data, which made estimation a challenge using maximum likelihood estimation.  Other authors, for example, Basu et al., used a Bayesian approach to fit the GFLP model to masked failure data from an engineering system \cite{basu}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.

\subsection{Motivation}
The goal of this paper is to model, and compare, the lifetime distribution of the various hard drive brands using the GFLP lifetime model.  There are numerous features of the data, however, that make standard estimation approaches for this parametric model problematic.  The first issue is many hard drives have been running for quite some time before entering the sample.  Even hard drives starting after 2013 appear in the data after hundreds of hours of operation.  When an observation is at risk before entering the study, this is known as left truncation.  Also, although we have four years of data, hard drives rarely fail so many units are still in service.  Thus, we have both left truncation and right censoring.  The other challenge is each hard drive model varies in number units in operation and total time on test.  Some models have hundreds of units in service; others have as few as two.  Combined with the truncation and censoring this makes stable maximum likelihood (ML) estimates difficult to obtain for many drive models. \\

Another issue is product populations often contain a mixture of defective and nondefective units.  The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.  Ignoring this hazard structure, which is known to exist for computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.  The general limited failure population model (GFLP) combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.  Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are unknown, which presents challenges for parameter identification in the GFLP model, as discussed in Chan and Meeker (1999).  Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.  With over 60 brand drive models in testing a hierarchical framework is advantageous as it pools information from across brands to get more precise estimates of the individual parameters.  This is especially helpful for the Backblaze data as there are many drive models with as few as one or two failures.

\subsection{Overview}
The structure of the paper is as follows.  Section 1 summarizes the Backblaze data and the unique modeling challenges it presents.  Section 2 introduces the GFLP model as a mixture of two Weibull distributions.  In Section 3 we fit multiple specifications of the GFLP model to best model the Backblaze data.  We also discuss computational and modeling issues that suggests a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.  In Section 4 we compare hard drive brand models based on our final GFLP model.  Finally, in Section 5we review the GFLP model and propose future extensions.

>>>>

\subsection{Background}
Often products can fail due to more than one failure mode.  For example, there are many parts in a washing machine that can break causing the whole machine to fail.  Decorative tree lighting can fail if one bulb in the chain burns out.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \cite{nelson}.  A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.  Thus, the overall reliability of a unit is modeled using the product rule across all $s$ risks.  The GFLP model is a special case of the $s$ independent competing risks model with 2 modes of failure: defects and wearout.  The GFLP model was first introduced to describe the early and late failure modes of CB radios \cite{chan}.  Chan and Meeker used maximum likelihood (ML) estimation to estimate the model parameters, and developed likelihood based approximation methods to construct confidence intervals for functions of model parameters.  Other authors, for example, Basu et al., have employed a Bayesian approach to fit the GFLP model to masked failure data from an engineering system \cite{basu}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.  There has not been, however, any extensions of the GFLP model to a hierarchical framework.

\subsection{Motivation}
The goal of this paper is to model, and compare, the lifetime distribution of the various hard drive brands using the GFLP lifetime model.  There are numerous features of the data, however, that make standard estimation approaches for this parametric model problematic.  The first issue is many hard drives have been running for quite some time before entering the sample.  Even hard drives starting after 2013 appear in the data after hundreds of hours of operation.  When an observation is at risk before entering the study, this is known as left truncation.  Also, although we have four years of data, hard drives rarely fail so many units are still in service.  Thus, we have both left truncation and right censoring.  The other challenge is each hard drive model varies in number units in operation and total time on test.  Some models have hundreds of units in service; others have as few as two.  Combined with the truncation and censoring this makes stable ML estimates difficult to obtain for many drive models. \\

Another issue is product populations often contain a mixture of defective and nondefective units.  The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.  Ignoring this hazard structure, which is known to exist for computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.  The general limited failure population model (GFLP) combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.  Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are unknown, which presents challenges for parameter identification in the GFLP model, as discussed in Chan and Meeker (1999).  Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.  With over 60 brand drive models in testing a hierarchical framework is advantageous as it pools information from across brands to get more precise estimates of the individual parameters.  This is especially helpful for the Backblaze data as there are many drive models with one or two failures.

\subsection{Overview}
The structure of the paper is as follows.  Section 1 summarizes the Backblaze data and the unique modeling challenges it presents.  Section 2 introduces the GFLP model as a mixture of two Weibull distributions.  In Section 3 we fit multiple specifications of the GFLP model to best describe the Backblaze data.  We also discuss computational and modeling issues that suggest a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.  In Section 4 we compare hard drive brand models based on our final GFLP model.  Finally, in Section 5, we review the GFLP model and propose future extensions.
